---
title: "Nested Bayes Analysis"
author: "b.a.e"
format: html
editor: visual
---

This is the Bayesian analysis of Jinyu's data. We fit it with brms for legibility, the idea is to model the interaction between valences and stimuli with random slopes and intercepts for individuals. The variables of interest are "Intervention", "Pic_valence", "Subject" and "Valence_post" as the outcome. "Valence_post" needs to be standardised. First we subset the data to a neat dataframe for analysis.

```{r}
# Loading data
library(lme4)
library(tidyverse)
library(readxl)
library(haven)
df_valence <- read_excel("data/nested_Exp1_valence_imme_ratings_diff.xlsx")

# Standardising valence 
zv <- c(scale(df_valence$Valence_post,center = T,scale = T))
# Extracting within person observations
id <- df_valence$Subject
# Extracting valence factor
valence <- df_valence$Pic_Valence
# Extracting intervention factor
intervention <- df_valence$Intervention
# Extracting image
image <- df_valence$Image
#Making interaction 
int_vl <- interaction(df_valence$Intervention, df_valence$Pic_Valence)

df <- data.frame(id = factor(id), 
                 image = factor(image),
                 valence = factor(valence), 
                 intervention = factor(intervention), 
                 int_vl,
                 zv)
```

# Frequentist analyses

```{r}
summary(lmer(zv ~ valence + (1|id) , data=df_valence, REML=T))
summary(lmer(zv ~ valence + intervention + (1|id) , data=df_valence, REML=T))

# Base model
base_model <- lmer(zv ~ valence * intervention  + (1|id), data=df_valence, REML=T)

# Suggested by other papers
suggested_model <- lmer(zv ~ valence * intervention + (1|id) + (1|image), data=df_valence, REML=T)

# Nested design
nested_model <- lmer(zv ~ valence * intervention + (1|id) + (1|image:intervention), data=df_valence, REML=T) # Model 5 for Alfons

# Start the issues
summary(lmer(zv ~ valence * intervention  + (1|id) + (1|valence:intervention), data=df_valence, REML = T)) # Model 6 for Alfons 

# Tested models 
summary(lmer(zv ~ valence * intervention + (1|id) + (1|valence:intervention), data= df_valence, REML = T))

summary(lmer(zv ~ valence * intervention + (1|id/valence:intervention) , data=df_valence, REML=T))

summary(lmer(zv ~ valence * intervention + (1|valence:intervention) + (1 + valence/intervention), data=df_valence, REML=T))

summary(lmer(zv ~ valence * intervention + (image|id/valence:intervention), data=df_valence, REML=T))

summary(lmer(zv ~ valence * intervention + (1+ valence*intervention|id) + (1/valence:intervention:image), data=df_valence, REML=T))

summary(lmer(zv ~ valence * intervention + (1+ valence:intervention|id) + (1/valence:intervention:image), data=df_valence, REML=T))

summary(lmer(zv ~ valence * intervention + (1+valence+intervention|id), data=df_valence, REML=T))

summary(aov(zv ~ valence * intervention, data = df_valence))
```

```{r}
# The last model without issues
model_rand_slope <- lmer(zv ~ valence * intervention  + (1 + valence|id) + (1|intervention:image), data = df_valence, REML = T)

# Still good
model3 <- lmer(zv ~ valence * intervention  + (1|id/valence:intervention), data=df_valence, REML = T)

# It works!
model4 <- lmer(zv ~ valence * intervention + (1|int_vl:image), data=df_valence, REML = T)

# Still works; including id
model5 <- lmer(zv ~ valence * intervention + (1|id) + (1|int_vl:image), data=df_valence, REML = T) # Same as model3

```

No random slope for this model, it is too complex and it renders warnings and errors.

$R^2$ Marginal: Only fixed effects.

```{r}
############################ FINAL MODELS ######################################
# ANOVA
summary(aov(zv ~ valence * intervention, data = df_valence))

# Base model
base_model <- lmer(zv ~ valence * intervention  + (1|id), data=df_valence, REML=T)

# Suggested by other papers
suggested_model <- lmer(zv ~ valence * intervention + (1|id) + (1|image), data=df_valence, REML=T)

# The last model without issues
model_rand_slope <- lmer(zv ~ valence * intervention  + (1 + valence|id) + (1|intervention:image), data = df_valence, REML = T)


anova(suggested_model, model4)
```

# Bayesian analyses

The target model is a mixed effects regression with random slopes and intercepts for each individual, specifically targeting the interaction between valences and intervention. Since the data is massive, we try out the terget model on a smaller subset. We specify the formula:

```{r}
df_sub <- df[1:500,]

library(brms)
formula <- (
    zv ~ valence * intervention
    + (1 + valence:intervention | id)  # participant random intercept
    + (1 | valence/intervention)       # stimuli nested in valence
)
```

We can then specify the priors, going for weakly informative ones initially.

```{r}
priors <- prior(
  # Fixed effects — outcome is standardised.
  prior(normal(0, 1),   class = b), # Slope for id
  prior(normal(0, 1),   class = Intercept), # Intercept for id

  # SDs for participant random effects (intercept + interaction slope)
  prior(exponential(1),  class = sd, group = participant),

  # SDs for intervention in valence (negative/control)
  prior(exponential(1),  class = sd, group = valence),
  prior(exponential(1),  class = sd, group = valence:intervention),

  # Correlation between participant intercept and interaction slope
  prior(lkj(2), class = cor,  group = participant),

  # Residual SD
  prior(exponential(1),  class = sigma)
)
```

We fit the model

```{r}
fit <- brm(
  formula = formula,
  data    = df,
  family  = gaussian(),
  chains  = 4,
  iter    = 4000, 
  control = list(adapt_delta = 0.95),
  seed    = 42 # Answer to the meaning of life, the universe, and everything.
)

summary(fit)
plot(fit)
pairs(fit)
```
